Re-examining cross-cultural similarity judgments using language statistics

### 2023-02-02 code/4_paper/manuscript/manuscript.tex
final pdf

\begin{abstract}
Is ``cow'' more closely related to ``grass'' or to ``chicken''? Speakers
of different languages judge similarity in this context differently, but
why? One possibility is that cultures co-varying with these languages
induce differences in conceptualizations of similarity. Specifically,
East Asian cultures may promote reasoning about thematic similarity, by
which cow and grass are more related, whereas Western cultures may bias
judgments toward taxonomic relations, like cow and chicken. We measure
similarity judgments across the US, China, and Vietnam and replicate
US-China cultural differences with Vietnam intermediate. We also show
that word embedding models (fastText models for each language) are
related to judgments within each country, suggesting a possible
alternative interpretation of cross cultural differences. Perhaps
notions of similarity are similar across contexts, but the statistics of
the linguistic environment vary.

@@

Additionally, language, culture, cognition, and individual experiences
are intertwined in complex causal relationships. In this study, we
measure language and its relation to cross-cultural differences in
categorization, but these relations test only the plausibility of a
language-based account; they cannot establish the direction of
causality.

Ji et al. (2004) established that culture-aligned differences in this
paradigm exist, even when the test language is held constant, concluding
that ``it is culture (independent of the testing language) that led to
different grouping styles'' in their study. Our data provide a
cautionary note to this conclusion, suggesting that while cultural
differences in similarity judgment exists, we might get better traction
on modelling cross-cultural variation through environmental statistics
such as lexical co-occurrences.

There are still many open questions for this account, however. Our
operationalization of environmental statistics with word embeddings
likely captures both language inputs and culture-specific
conceptualization of similarity. Further investigation of the semantic
spaces captured by cross-cultural word embedding models is required.
Future work should also aim to provide a more specific computational
account of how lexical co-occurrence might guide categorization
preference beyond the simple proportion-of-similarity model tested here,
and investigate other individual- or triad-specific factors that might
influence how environmental statistics affect similarity judgment.

>>>>

\begin{abstract}
Is ``cow'' more closely related to ``grass'' or to ``chicken''? Speakers
of different languages judge similarity in this context differently, but
why? One possibility is that cultures co-varying with these languages
induce differences in conceptualizations of similarity. Specifically,
East Asian cultures may promote reasoning about thematic similarity, by
which cow and grass are more related, whereas Western cultures may bias
judgments toward taxonomic relations, like cow and chicken. We measure
similarity judgments across the US, China, and Vietnam and replicate
US-China differences, but do not find that responding in Vietnam
patterns with China. Instead, similarity judgments in Vietnam are
intermediate between the US and China. We also show that word embedding
models (fastText models for each language) are related to judgments
within each country, suggesting a possible alternative interpretation of
cross cultural differences. Perhaps notions of similarity are similar
across contexts, but the statistics of the linguistic environment vary.

@@

Additionally, language, culture, cognition, and individual experiences
are intertwined in complex causal relationships. In this study, we
measure language and its relation to cross-cultural differences in
similarity judgments, but these relations test only the plausibility of
a language-based account; they cannot establish the direction of
causality.

Ji et al. (2004) established that culture-aligned differences in this
paradigm exist, even when the test language is held constant, concluding
that ``it is culture (independent of the testing language) that led to
different grouping styles'' in their study. Our data provide a
cautionary note to this conclusion, suggesting that while cultural
differences in similarity judgments exist, we might get better traction
on modelling cross-cultural variation through environmental statistics
such as lexical co-occurrence.

There are still many open questions for this account, however. Our
operationalization of environmental statistics with word embeddings
likely captures both language inputs and culture-specific
conceptualization of similarity. Further investigation of the semantic
spaces captured by cross-cultural word embedding models is required.
Future work should also aim to provide a more specific computational
account of how lexical co-occurrence might guide categorization
preference beyond the simple proportion-of-similarity model tested here,
and investigate other person- or triad-specific factors that might
influence how environmental statistics affect similarity judgments.
